---
title: "BTC model"
author: "Calin Uioreanu"
date: "February 26, 2018"
output: html_document
#output:
#  md_document:
#    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## BTC Modeling

This is an attempt to model the BTC price using previous days data.

```{r requirements, echo=FALSE,include=FALSE, cache=FALSE}
library(Quandl)
library(ggplot2)
library(dplyr)
library(e1071)
library(caret)
library(randomForest)
#library(xtable)
```

## Retrieving BTC data

We use the Quandl API to retrieve historical BTC prices, filter >= 2014

```{r data, echo=TRUE}
btc <- Quandl("BCHARTS/BTCDEEUR")
btc <- subset(btc, Date>='2014-01-01')
dim(btc)
```


## BTC trends

the daily weighted BTC price

```{r chart, echo=FALSE}
ggplot(btc, aes(x = Date, y = `Weighted Price`)) + 
  geom_point() + 
  theme_minimal() + 
  geom_smooth(method = "loess")
```

## BTC linear regression

a simple linear regression model to see how the prices correlate

```{r lm.simple, echo=FALSE}
lm.model <- lm (`Weighted Price` ~ Open + High + Low + Close, data = btc)
summary(lm.model)
```

So intra-day data is highly correlated. Preparing prediction using historical data:

## Data preparation

historical data via lag

```{r lag_data, echo=TRUE}
ds <- btc %>%
    mutate(Price = `Weighted Price`) %>%
    dplyr::select(Date, Price) %>%
    arrange(Date) %>%
    mutate(Grow_0 = Price > lag(Price, 1), 
           Grow_1 = lag(Price, 1)>lag(Price, 2), 
           Grow_2 = lag(Price, 2)>lag(Price, 3), 
           Grow_3 = lag(Price, 3)>lag(Price, 4), 
           Grow_4 = lag(Price, 4)>lag(Price, 5), 
           Grow_5 = lag(Price, 5)>lag(Price, 6), 
           Grow_6 = lag(Price, 6)>lag(Price, 7), 
           Grow_5 = lag(Price, 7)>lag(Price, 8), 
           Grow_1w = lag(Price, 1)>lag(Price, 8), 
           Grow_2w = lag(Price, 1)>lag(Price, 15)) %>%
    dplyr::select(-Date, - Price)
  ds <- ds[complete.cases(ds),]
  head(ds)
```
## BTC Model using Naive Bayes

a logistic regression model to see how the today's trend correlates to historical.

```{r btc_model, echo=TRUE}
x = ds[, -1]
  y = as.factor(ds$Grow_0)
  library(e1071)
  model = train(
    x,y,model='nb',
    trControl=trainControl(method='cv',number=10), metric="Accuracy")
  model
```
  
Tools: caret package, randomForest and 10 fold Cross-validation
