---
title: "BTC model"
author: "Calin Uioreanu"
date: "February 26, 2018"
output: html_document
#output:
#  md_document:
#    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## BTC Modeling

This is an attempt to model the BTC price using previous days data.

```{r requirements, echo=FALSE,include=FALSE, cache=FALSE}
library(Quandl)
library(ggplot2)
library(dplyr)
library(e1071)
library(caret)
library(randomForest)
library(e1071)
#library(xtable)
```

## Retrieving BTC data

We use the Quandl API to retrieve historical BTC prices, filter >= 2014

```{r data, echo=TRUE}
btc <- Quandl("BCHARTS/BTCDEEUR")
btc <- subset(btc, Date>='2014-01-01')
dim(btc)
```


## BTC trends

the daily weighted BTC price

```{r chart, echo=FALSE}
ggplot(btc, aes(x = Date, y = `Weighted Price`)) + 
  geom_point() + 
  theme_minimal() + 
  geom_smooth(method = "loess")
```

## Dataset correlation coefficients

a simple correlationchart, showing the high degree of correlation between intra-day data

```{r lm.simple, echo=FALSE}
corrplot::corrplot(
  cor(btc[, -1]),type="upper", method="pie", order = "hclust", addrect = 3)
```

# Data Preparation
### Preparing prediction using historical data

historical data via lag

```{r lag_data, echo=TRUE}
ds <- btc %>%
    mutate(Price = `Weighted Price`) %>%
    dplyr::select(Date, Price) %>%
    arrange(Date) %>%
    mutate(Grow_0 = Price > lag(Price, 1), 
           Grow_1 = lag(Price, 1)>lag(Price, 2), 
           Grow_2 = lag(Price, 2)>lag(Price, 3), 
           Grow_3 = lag(Price, 3)>lag(Price, 4), 
           Grow_4 = lag(Price, 4)>lag(Price, 5), 
           Grow_5 = lag(Price, 5)>lag(Price, 6), 
           Grow_6 = lag(Price, 6)>lag(Price, 7), 
           Grow_5 = lag(Price, 7)>lag(Price, 8), 
           Grow_1w = lag(Price, 1)>lag(Price, 8), 
           Grow_2w = lag(Price, 1)>lag(Price, 15)) %>%
    dplyr::select(-Date, - Price)
rownames(ds) <- btc$Date

  ds <- ds[complete.cases(ds),]
  head(ds)
  x = ds[, -1]
  y = as.factor(ds$Grow_0)
```

# Prediction comparison
### BTC Modeling using several ML algos

a logistic regression model to see how today's trend can be predicted based on historical data

Tools: 4 popular ML method x 30 results ( 3 repeats of 10-fold cross validation)

Methods: RF, lvw, svm, gbm

```{r btc_model, echo=TRUE}

control = trainControl(method='cv',number=10, repeats=3)

set.seed(123)
model.rf  = train(x,y,model='rf' ,trControl=control, metric="Accuracy")
model.lvq = train(x,y,model='lvq',trControl=control, metric="Accuracy")
model.gbm = train(x,y,model='gbm',trControl=control, metric="Accuracy", verbose=FALSE)
model.svm = train(x,y,model='svm',trControl=control, metric="Accuracy")

# collect resamples
results <- resamples(list(RF = model.rf, LVQ=model.lvq, GBM=model.gbm, SVM=model.svm))
# summarize the distributions
summary(results)
  
```

All methods perform generally well, but Gradient Boosting Machine has the lowest median accuracy, same a low Kappa. SVM has also the widest range:

```{r btc_model_charts, echo=TRUE}
# boxplots of results
bwplot(results)
# dot plots of results
dotplot(results)
```

# Best method: Learning Vector Quantization lvq
### mean Accuracy 59,2%
### mean Kappa: 16,5%
Kappa = the amount of agreement correct by the agreement expected by chance
